Towards Verifying the Bitcoin-S Library

### 2019-06-24 paper/paper.tex
more changes

\subsection{Creation of a Transaction}

@@

\subsection{Validation of a Transaction}

@@

\paragraph{Stainless integration problems.}

\todo{Kai: check the text}

@@

\paragraph{Extracting the relevant code.}
The extracted code has more than 1500 lines.  After running Stainless
on it, it still throws a huge bunch of errors about what Stainless can
not reason about. After fixing some of those errors, there appear new
ones.  So this would require changing nearly everything of the
extracted code.

Let's adjust the property from the \nameref{property_1} to the
\nameref{property_2}, because there is not enough time to fix all
these errors and write the verification.  This is way less interesting
to verify but gives still many insights in what needs to be done to
verify some parts of Bitcoin-S and other source code.

@@

\subsection{Extracting the relevant Code}

@@

\subsection{Result}

Finally, everything is green and correctly verified. \todo{kai special chars}
% \begin{lstlisting}[style=scala]
% [Warning ] The Z3 native interface is not available. Falling back onto smt-z3.
% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 1))' VC for one @69:18...
% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
% [  Info  ] Cache miss: 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
% [  Info  ]  - Now solving 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5...
% [  Info  ]  - Result for 'precond. (call apply(Int64(), bigInt))' VC for apply @52:5:
% [  Info  ]  => VALID
% [  Info  ]  - Checking cache: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
% [  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
% [  Info  ] Cache hit: 'precond. (call +(underlying(satoshis(thiss)), underlyin ...)' VC for + @14:14...
% [  Info  ] Cache miss: 'postcondition' VC for + @12:3...
% [  Info  ]  - Now solving 'postcondition' VC for + @12:3...
% [  Info  ]  - Result for 'postcondition' VC for + @12:3:
% [  Info  ]  => VALID
% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
% [  Info  ] Cache hit: 'precond. (call apply(Int64(), -9223372036854775808))' VC for min @71:18...
% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
% [  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 9223372036854775807))' VC for max @72:18...
% [  Info  ]  - Checking cache: 'adt invariant' VC for apply @79:5...
% [  Info  ] Cache hit: 'adt invariant' VC for apply @79:5...
% [  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
% [  Info  ] Cache hit: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @22:11...
% [  Info  ]  - Checking cache: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
% [  Info  ] Cache hit: 'precond. (call apply(Int64(), 0))' VC for zero @68:19...
% [  Info  ]   ┌───────────────────┐
% [  Info  ] ╔═╡ stainless summary ╞══════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╗
% [  Info  ] ║ └───────────────────┘                                                                                                                      ║
% [  Info  ] ║ +      postcondition                                                 valid             U:smt-z3  currency/CurrencyUnits.scala:12:3   0.643 ║
% [  Info  ] ║ +      precond. (call +(underlying(satoshis(thiss)), underlyin ...)  valid from cache            currency/CurrencyUnits.scala:14:14  0.097 ║
% [  Info  ] ║ +      precond. (call checkResult(thiss, underlying(thiss) + u ...)  valid from cache            number/NumberType.scala:22:11       0.015 ║
% [  Info  ] ║ apply  adt invariant                                                 valid from cache            currency/CurrencyUnits.scala:34:39  0.005 ║
% [  Info  ] ║ apply  precond. (call apply(Int64(), bigInt))                        valid             U:smt-z3  number/NumberType.scala:52:5        0.187 ║
% [  Info  ] ║ apply  adt invariant                                                 valid from cache            number/NumberType.scala:79:5        0.017 ║
% [  Info  ] ║ max    precond. (call apply(Int64(), 9223372036854775807))           valid from cache            number/NumberType.scala:72:18       0.001 ║
% [  Info  ] ║ min    precond. (call apply(Int64(), -9223372036854775808))          valid from cache            number/NumberType.scala:71:18       0.001 ║
% [  Info  ] ║ one    precond. (call apply(Int64(), 1))                             valid from cache            number/NumberType.scala:69:18       0.915 ║
% [  Info  ] ║ zero   precond. (call apply(Int64(), 0))                             valid from cache            number/NumberType.scala:68:19       0.001 ║
% [  Info  ] ╟┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄┄╢
% [  Info  ] ║ total: 10   valid: 10   (8 from cache) invalid: 0    unknown: 0    time:   1.882                                                           ║
% [  Info  ] ╚════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════════╝
% [  Info  ] Shutting down executor service.
% \end{lstlisting}

% \begin{figure}
% 	\centering
% 		\includegraphics[width=\textwidth]{final_verify_output}
% 	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
% \end{figure}

@@

But is it really the original code that we verified?
And why did we have to change that much?
This and other questions will be answered in the next chapter.

\section{Practical Challenges with Stainless}
\label{chap:appendix_arb}

@@

\subsection{sbt vs JAR}

@@

However, currently there are some drawbacks.  For example the sbt
plugin does not always report errors.  We will see more about that
shortly.

@@

\appendix
\section{Code Transformations}

\paragraph{Project Setup}
\begin{itemize}
\item clone our repo
\item contains a stainless jar
\item call the script blahblah
\end{itemize}

>>>>

\paragraph{Creating a Transaction}

@@

\paragraph{Validating a Transaction}

@@

\paragraph{Trying Stainless on the entire project.}

\begin{itemize}
\item result sbt: no output
\item result jar: 
\end{itemize}

@@

\paragraph{Putting aside the No-Inflation Property.} Naively trying
Stainless on the entire bitcoin-s codebase results in either no output
(with sbt) or many errors (with jar). So we extract the relevant code
to only verify that. However, the extracted code has more than 1500
lines and liberally uses Scala features outside of the supported
fragment.  We tried to transform the code into the supported fragment,
but realize that a better approach is to first verify a simpler
property with less code involved and then turn back to the
no-inflation property with more experience. So we turn to the
addition-with-zero property in the next section. \todo{explain}

@@

\paragraph{Extracting the relevant Code}

@@

\paragraph{Transforming the Code}

\todo{kai todo}

\paragraph{Result}

Finally, everything is green and correctly verified. 
\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{final_verify_output}
	\caption{Output of Stainless verification for addition with 0 of Bitcoin-S-Cores CurrencyUnit}
\end{figure}

@@

\appendix
\section{Project Setup}

@@

\todo{nur über stainless-sbt-plugin reden damit man weiss wir haben es versucht}

\subsection{stainless-sbt-plugin vs JAR}

@@

However, currently there are some drawbacks.  For example the sbt
plugin does not always report errors.

We use the jar for everything.

@@

\begin{itemize}
\item clone our repo
\item jar vs sbt, you can use both
\item contains a stainless jar
\item call the script blahblah
\end{itemize}

\section{Code Transformations}

These are the code transformations needed to Verify the Addition-with-Zero Property.
